\documentclass{IEEEtran}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
\else
   \usepackage[dvips]{graphicx}
\fi

\ifCLASSOPTIONcompsoc
  \usepackage[caption=false, font=normalsize, labelfont=sf, textfont=sf]{subfig}
\else
  \usepackage[caption=false, font=footnotesize]{subfig}
\fi

\usepackage{textcomp}
\usepackage{nicefrac}
\usepackage{siunitx}
\usepackage{fancyref}

\usepackage[style=ieee, doi=false, isbn=false, url=false, maxbibnames=1, minbibnames=1, maxcitenames=1, mincitenames=1, backend=biber, defernumbers=false]{biblatex}
\addbibresource{./references.bib}

\AtEveryBibitem{\clearfield{month}}
\AtEveryBibitem{\clearfield{day}}
\AtEveryBibitem{\clearfield{volume}}
\AtEveryBibitem{\clearfield{issue}}
\AtEveryBibitem{\clearfield{pages}}
\AtEveryBibitem{\clearfield{number}}
\AtEveryBibitem{\clearfield{title}}
\AtEveryBibitem{\clearfield{isbn}}
\AtEveryBibitem{\clearfield{keywords}}
\AtEveryBibitem{\clearfield{issn}}
\AtEveryBibitem{\clearfield{journal}}

\usepackage[activate={true, nocompatibility}, final, tracking=true, kerning=true, spacing=true, factor=1100, stretch=10, shrink=10]{microtype}
\linespread{0.9}

\input{glossary}

% \markboth{IEEE TRANSACTIONS ON NUCLEAR SCIENCE, VOL. XX, NO. XX, XXXX 2020}
% {Author \MakeLowercase{\textit{et al.}}: Preparation of Papers for Review by the \textsc{IEEE Transactions on Nuclear  Science} \newline (May 2020)}

\usepackage{lipsum}

\begin{document}
    \title{
        \vspace{-0.75cm}
        
        A Bayesian Neural Network-Based Method for the Extraction of a Metabolite Corrected Arterial Input Function from Dynamic [$^{11}$C]PBR28 PET 
    }
    
    \author{
        \vspace{-0.25cm}
        
        Alexander~C.~Whitehead$^{*}$~\IEEEmembership{Student~Member,~IEEE},
        Ludovica~Brusaferri$^{*}$,
        Lucia~Maccioni,
        Matteo~Ferrante,
        Marianna~Inglese,
        Zeynab~Alshelh,
        Mattia~Veronese,
        Nicola~Toschi,
        Jodi~Gilman,
        Kris~Thielemans,
        and~Marco~L.~Loggia
    
        \vspace{-0.75cm}
    
        \thanks{
            \scriptsize
            This work was funded by GE Healthcare, the NIHR UCLH Biomedical Research Centre, the UCL EPSRC Centre for Doctoral Training in Intelligent, Integrated Imaging in Healthcare (i4health) grant (EP/L016478/1), the Open Source Imaging Consortium (OSIC), the NIH grants R01-NS094306-01A1, R01-NS095937-01A1 and R01-DA047088-01, the Italian Ministry of University and Research (MUR), National Recovery and Resilience Plan(NRRP), project MNESYS (PE0000006)(to NT) – A Multiscale integrated approach to the study of the nervous system in health and disease (DN. 1553 11.10.2022), National Center for HPC, BIG DATA AND QUANTUM COMPUTING (Project no. CN00000013 CN1), the PNRR National Grant DIGITAL LIFELONG PREVENTION (Project no PNC0000002DARE), and by Wellcome Trust Digital Award (no. 215747/Z/19/Z).
        }
        \thanks{
            \scriptsize
            Alexander~C.~Whitehead is with the Department of Computer Science, University College London, London, UK. Ludovica~Brusaferri, Zeynab~Alshelh, Jodi~Gilman, Nicola~Toschi, and Marco~L.~Loggia are with Athinoula A. Martinos Center for Biomedical Imaging, Harvard Medical School, Boston, MA, US. Lucia~Maccioni and Mattia~Veronese are with the Department of Information Engineering, University of Padua, Padua, Italy and Neuroimaging Department, IoPPN, King’s College London, London, UK. Matteo~Ferrante, Marianna~Inglese, and Nicola~Toschi are with the Department of Biomedicine and Prevention, University of Rome Tor Vergata, Rome, Italy.
        }
    }
    
    \pagestyle{plain}
    \pagenumbering{gobble}
    
    \maketitle
    
    \begin{abstract}
        In Positron Emission Tomography (PET), arterial sampling and metabolite correction are prerequisites for the gold-standard measurement of values like the volume of distribution ($\mathrm{V_T}$), often necessary for the full quantification of radioligand binding. However, the invasiveness and technical demands of these procedures limit their application in both research and clinical PET studies. Machine learning approaches have been explored to predict $\mathrm{V_T}$ from PET images, but their integration in clinical routine is limited by their lack of transparency or thorough evaluation. Here we propose a Bayesian Neural Network to estimate the arterial input function (AIF), while also outputting its prediction uncertainty, 1) directly from the entire dynamic PET images (NN-AEIF), 2) from an image-derived input function (IDIF) (NN-IDIF) and, as a sensitivity measure, 3) from the un-corrected plasma curve (NN-AIF). All methods, applied on [$^{11}$C]PBR28 PET data, were compared to the metabolite-corrected AIF in terms of $\mathrm{V_T}$, and the prediction uncertainty was assessed in terms of normalised coefficient of variance (nCV). Overall, both NN-AEIF and NN-AIF were able to accurately predict $\mathrm{V_T}$, outperforming the other methods, with NN-AEIF showing the lowest nCV.
    \end{abstract}
    
    % \begin{IEEEkeywords}
      %   Arterial Input Function Estimation, Signal Extraction, Dynamic PET, Machine Learning, Bayesian Neural Networks
     % \end{IEEEkeywords}
    
    \input{introduction}
    \input{methods}
    \input{results}
    \input{discussion_and_conclusion}
    
    \vspace{-0.5cm}
    
    \AtNextBibliography{
        \scriptsize
    }
    \printbibliography
\end{document}
